---
params:
  sizeCut: 130
  includeDMEK: TRUE
  includeDSAEK: TRUE
  includeUT_DSAEK: TRUE
  
header-includes:
  - \newcommand\latexcode[1]{#1}
title: "Correlation analysis of Surgery duration and ECD"
author: "Alexis Derumigny"
date: "`r Sys.Date()`"
output: 
  bookdown::word_document2:
    toc: true
    fig_caption: true

knit: |
  (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = paste0(
        xfun::sans_ext(input), '-', Sys.Date(), '.docx'
      ),
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# Loading packages
library(readxl)
library(tidyverse)
library(ggforce)
library(ade4)
library(adegraphics)

# Loading cleaning functions
myFolder = here::here()
source("functions_cleaning.R")
```

# Introduction

```{r, echo=FALSE, results='asis', label="preparing the data frame"}
result = knitr::knit_expand(file = "preparing_dataframe.Rmd")

if (interactive()){
  result_ = knitr::knit(
    text = result, envir = globalenv(), quiet = TRUE)
} else {
  result_ = knitr::knit_child(
    text = result, envir = globalenv(), quiet = TRUE)
}

cat(result_, sep = '\n')
```


```{r}
myVariables = c("ECD.preop", "ECD_M6", "ECD_Y1", "ECD_Y2", "surgery.duration")

X = totalData.frame |>
  filter(is.dmek == 1) |>
  select(all_of(myVariables)) |>
  as.data.frame() |>
  as.matrix.data.frame()

corMatrix = cor(X, use = "pairwise.complete.obs")
```


# Correlation matrix

Here is the correlation matrix of the 5 variables:

```{r}
corrplot::corrplot(corMatrix, method = "color", col = corrplot::COL2('PiYG'), 
                   addCoef.col = 'black')
```


And here is the corresponding matrix with the p-values associated to the test
that a given correlation is significantly different from 0 or not.

```{r}
testRes = corrplot::cor.mtest(X, conf.level = 0.95)
corrplot::corrplot(corMatrix, p.mat = testRes$p, insig = 'p-value', sig.level = -1, 
                   col = corrplot::COL2('PiYG'), method = "color",
                   number.digits = 4)
```


# Correlation scatterplot


On the following figure, you can see:

- the scatterplots of each pair of variables;
- the estimated correlations with the stars to show the significance (or lack thereof);
- as a bonus: the estimated density of each variable by kernel smoothing
(in the diagonal facets).

```{r, fig.height=8, fig.width=8}
GGally::ggpairs(totalData.frame[, myVariables],
                progress = FALSE # No progress bar
) + theme(panel.spacing = unit(0.18, "inches"))
```



```{r function for PCA}
# Removing unnecessary columns
Xnorm = matrix(nrow = nrow(X), ncol = ncol(X))
for (j in 1:ncol(X)){
  Xnorm[, j] = (X[, j] - mean(X[, j], na.rm = TRUE)) / sd(X[, j], na.rm = TRUE)
}


my_PCA <- function(cormat)
{
  # 1- Diagonalization
  res_pca = eigen(cormat)
  
  # 2- Scree plot
  plot_scree = ggplot(data=NULL,
                      aes(x=1:length(res_pca$values),
                          y=res_pca$values/sum(res_pca$values))) +
    # geom_point(size=2)+
    geom_col()+
    # labs(title="Scree plot of the scaled PCA") +
    xlab("Principal components") +
    ylab("Eigenvalue / Sum(eigenvalues)")
  
  # 3- Biplot 
  vecpropres_1 = res_pca$vectors[,1]
  vecpropres_2 = res_pca$vectors[,2]
  vecpropres_12 = cbind(vecpropres_1, vecpropres_2)
  obs_firstplane = Xnorm %*% vecpropres_12
  
  PCA12 <- ggplot() + geom_point(aes(x = obs_firstplane[,1],
                                     y = obs_firstplane[,2]), size = 1.5) +
    theme(text=element_text(size=14),
          axis.text = element_text(size=10) ) +
    theme(panel.grid.minor = element_line(size = 1)) +
    theme(legend.spacing.y = unit(0.01, "npc"),
          legend.text =
            element_text(margin = margin(3,0,3,0))) +
    xlab("PC1") + ylab("PC2")
  
  
  # 4- Correlation circle
  val_plus = res_pca$values * as.numeric(res_pca$values >= 0)
  
  cor.factors <- NULL
  for (j in 1:ncol(X)){
    rf <- sqrt(val_plus[j])*res_pca$vectors[,j]
    cor.factors <- cbind(cor.factors,rf)
  }
  rownames(cor.factors) <- colnames(X)
  colnames(cor.factors) <- paste("F",1:ncol(X),sep="")
  
  # 5- Factor loadings
  DataFrameLoadings = data.frame(
    Variable = factor(rep(colnames(X), times = 2),
                      levels = sort(colnames(X), decreasing = TRUE)),
    Loadings = c(cor.factors[,1], cor.factors[,2]),
    Factor = rep(c("PC1","PC2"), each = ncol(X))
  )
  PlotLoadings = ggplot(DataFrameLoadings) +
    geom_col(aes(x = Variable, y = Loadings)) +
    facet_grid(cols = vars(Factor)) +
    coord_flip()
  
  return( list(res_pca = res_pca, plot_scree = plot_scree, 
               PCA12_ellipses = PCA12, cor.factors = cor.factors,
               PlotLoadings = PlotLoadings) )
}
```



```{r, echo=FALSE, results='asis', label="all PCAs"}
out = list()
listPCA = list()

i = 1
listPCA[[i]] = my_PCA(cormat = corMatrix)

nameAnalysis.short = "Surgery duration and ECD"

nameSection = "# PCA for Surgery duration and ECD"

# nameScreePlot = paste("'Scree plot -", nameAnalysis.short, "'")

result = knitr::knit_expand(file = "multiple-PCA-section.Rmd")
out[[i]] = result
  

if (interactive()){
  res <- lapply(out, function(x) {
    knitr::knit(
      text = x, envir = globalenv(), quiet = TRUE
    )
  })
} else {
  res <- lapply(out, function(x) {
    knitr::knit_child(
      text = x, envir = globalenv(), quiet = TRUE
    )
  })
}
cat(unlist(res), sep = '\n')
```


