---
params:
  sizeCut: 130
  includeDMEK: TRUE
  includeDSAEK: TRUE
  includeUT_DSAEK: TRUE
  
header-includes:
  - \newcommand\latexcode[1]{#1}
title: "Correlation analysis of Surgery duration and ECD"
author: "Alexis Derumigny"
date: "`r Sys.Date()`"
output: 
  bookdown::word_document2:
    toc: true
    fig_caption: true

knit: |
  (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = paste0(
        xfun::sans_ext(input), '-', Sys.Date(), '.docx'
      ),
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# Loading packages
library(readxl)
library(tidyverse)
library(ggforce)
library(ade4)
library(adegraphics)

# Loading cleaning functions
myFolder = here::here()
source("functions_cleaning.R")
```

# Introduction

```{r, echo=FALSE, results='asis', label="preparing the data frame"}
result = knitr::knit_expand(file = "preparing_dataframe.Rmd")

if (interactive()){
  result_ = knitr::knit(
    text = result, envir = globalenv(), quiet = TRUE)
} else {
  result_ = knitr::knit_child(
    text = result, envir = globalenv(), quiet = TRUE)
}

cat(result_, sep = '\n')
```


```{r}
myVariables = c("ECD_M6", "ECD_Y1", "ECD_Y2", "surgery.duration")

X = totalData.frame |>
  filter(is.dmek == 1) |>
  select(all_of(myVariables)) |>
  as.data.frame() |>
  as.matrix.data.frame()

corMatrix = cor(X, use = "pairwise.complete.obs")
```


# Correlation matrix

Here is the correlation matrix of the 4 variables:

```{r}
corrplot::corrplot(corMatrix, method = "color", col = corrplot::COL2('PiYG'), 
                   addCoef.col = 'black')
```


And here is the corresponding matrix with the p-values associated to the test
that a given correlation is significantly different from 0 or not.

```{r}
testRes = corrplot::cor.mtest(X, conf.level = 0.95)
corrplot::corrplot(corMatrix, p.mat = testRes$p, insig = 'p-value', sig.level = -1, 
                   col = corrplot::COL2('PiYG'), method = "color",
                   number.digits = 4)
```


# Correlation scatterplot


On the following figure, you can see:

- the scatterplots of each pair of variables;
- the estimated correlations with the stars to show the significance (or lack thereof);
- as a bonus: the estimated density of each variable by kernel smoothing
(in the diagonal facets).

```{r, fig.height=8, fig.width=8}
GGally::ggpairs(X,
                progress = FALSE # No progress bar
) + theme(panel.spacing = unit(0.18, "inches"))
```




