---
params:
  sizeCut: 130
  includeDMEK: TRUE
  includeDSAEK: TRUE
  includeUT_DSAEK: TRUE
  
header-includes:
  - \newcommand\latexcode[1]{#1}
title: "Linear regressions for UDD paper"
author: "Alexis Derumigny"
date: "23/01/2023"
output: 
  bookdown::word_document2:
    toc: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# Loading packages
library(readxl)
library(tidyverse)
library(ggforce)
library(ade4)
library(adegraphics)

# Loading cleaning functions
myFolder = here::here()
source("functions_cleaning.R")
```

# Introduction

```{r, echo=FALSE, results='asis', label="preparing the data frame"}
result = knitr::knit_expand(file = "preparing_dataframe.Rmd")

result_ = knitr::knit_child(
    text = result, envir = globalenv(), quiet = TRUE)

cat(result_, sep = '\n')
```


```{r}
### Here are the list of available variables at this point
# sort(colnames(totalData.frame))
```


```{r}
# Filling out the NAs
totalData.frame$rebubbling[is.na(totalData.frame$rebubbling)] <- 0
totalData.frame$X2nd.rebubbling[is.na(totalData.frame$X2nd.rebubbling)] <- 0
```


# A few more visualizations

```{r}
ggplot(totalData.frame) + 
  geom_jitter(aes(x = patient.age, y = VA_logM12, color = method), width = 1, height = 0.1)
```

```{r}
ggplot(totalData.frame) + 
  geom_jitter(aes(x = ECD.preop, y = VA_logM12, color = method), width = 1, height = 0.1)
```

```{r}
ggplot(totalData.frame) + 
  geom_jitter(aes(x = X2nd.rebubbling, y = VA_logM12, color = method), width = 0.1, height = 0.1)
```

```{r}
ggplot(totalData.frame) + 
  geom_jitter(aes(x = IrvingGass, y = VA_logM12, color = method), width = 0.1, height = 0.1)
```


# Linear regressions


## A. 

(A)  Multiple linear regression to determine whether 12-month VA in all first eyes can be predicted by interesting factors on PCA ie keratoplasty type, indication, patient age, preoperative VA, preoperative ECD, Any-Rebubbling

```{r}
reg_A = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD.preop + rebubbling + IrvingGass)

summary(reg_A)
```

```{r}
reg_A = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD.preop + rebubbling + IrvingGass +
             is.dmek + is.ut_dsaek)

summary(reg_A)
```


## B. 

(B)   Repeat A substituting Any-Rebubbling with MR (AR and MR may have some collinearity)

```{r}
reg_B = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD.preop + X2nd.rebubbling + IrvingGass)
summary(reg_B)
```

```{r}
reg_B = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD.preop + X2nd.rebubbling + IrvingGass +
             is.dmek + is.ut_dsaek)

summary(reg_B)
```


## C.

(C & D) Repeat A & B substituting preoperative ECD with 24-month ECD

```{r}
reg_C = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y2 + rebubbling + IrvingGass)

summary(reg_C)
```

```{r}
reg_C = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y2 + rebubbling + IrvingGass +
             is.dmek + is.ut_dsaek)

summary(reg_C)
```

## D.

```{r}
reg_D = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y2 + X2nd.rebubbling + IrvingGass)

summary(reg_D)
```

```{r}
reg_D = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y2 + X2nd.rebubbling + IrvingGass +
             is.dmek + is.ut_dsaek)

summary(reg_D)
```


## E.

(E & F) Repeat A & B substituting preoperative ECD with 12-month ECD***

```{r}
reg_E = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y1 + rebubbling + IrvingGass)

summary(reg_E)
```

```{r}
reg_E = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y1 + rebubbling + IrvingGass +
             is.dmek + is.ut_dsaek)

summary(reg_E)
```


## F.

```{r}
reg_F = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y1 + X2nd.rebubbling + IrvingGass)

summary(reg_F)
```

```{r}
reg_F = lm(data = totalData.frame,
           formula = VA_logM12 ~ patient.age + ECD_Y1 + X2nd.rebubbling + IrvingGass +
             is.dmek + is.ut_dsaek)

summary(reg_F)
```

